---
title: "Computational Musicology- Final Portfolio"
author: "Alihan Ince"
date: ""
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme: "default"

---
Comparing French and Dutch Raps 
=========================================

Column 
-------------------------------------
### Exploring the Differences and Similarities Between Two Cultural Forms 

#### Introduction: The Corpus and the Purpose of the Study
[My corpus](https://open.spotify.com/playlist/7KtmJXjsDmw8WEdtyL28L3?si=0448a854bfd84eb5) consists of 300 tracks by the following French and Dutch rap artists and singers: Alonzo, Jul, PNL, Gims (French) and Lijpe, Boef, Sevn Alias, Josylvio (Dutch). So the corpus consists of both French and Dutch hip-hop music (raps).   
I deliberately chose this corpus because I often like listening to French and Dutch rappers myself.  
In terms of style and approach, French and Dutch rap are different, but both are characterized by their ability to engage with the issues and experiences of their respective communities, and both have made significant contributions to the development of hip-hop as a global cultural form. It is therefore worth researching rap music of both languages where we will look at the differences and similarities between the two.  
  
#### Differences and similairities between French and Dutch Rap
The natural groups are French rappers (Alonzo, Jul, PNL, Gims) and Dutch rappers (Lijpe, Boef, Sevn Alias, Josylvio).   
French rap is more characterized by its strong tradition of chanson (which is a style of song that emphasizes vocal performance and poetic lyrics) and French rappers often incorporate African and Carribean music into their work, while Dutch rappers incorporate elements of gabber and house into their work.  
Both groups produce rap which means that they both use a very fast and energetic flow as well as techniques such as Onomatopoeaia (sue of words that imitate sounds, such as "boom" or "snap").   
However, I am curious to see what the analyses will show. Each rapper obviously has his own way of rapping and, of course, does not always use the same technique. It is therefore important to generalise as much as possible when comparing raps in two languages. Moreover, I tried to keep the rappers as representative as possible.   
  
#### Selection of Rappers and Genres 
The chosen rappers produce raps in French and Dutch. So to make a comparison or look at differences between Dutch and French raps, I felt it was enough to choose two Dutch and two French artists.   
The corpus consist of raps and so it belongs to "Hip hop music" or "rap music". I think the tracks will cover the genre well since they are all raps. 
The only gap maybe is that tracks by rapper Alonzo are gangsta rap in addition to hip-hop. Furthermore, Jul also makes hip-house music and tracks by Boef and Lijpe belong to nederhop and street rap. ultimately, they do fall under the umbrella genre "Hip hop".  
  
#### Sample Tracks
Some typical typical sample tracks in my corpus are Ciao La France (Alonzo), Namek (Jul) and Hoofdpijn (Lijpe)   

Column 
-------------------------------------
### The Corpus 
```{=html}
<iframe src="https://open.spotify.com/embed/playlist/7KtmJXjsDmw8WEdtyL28L3?utm_source=generator&theme=0" width="100%" height="100%" frameBorder="0" allowtransparency="true" allow="encrypted-media" data-external="1"></iframe>
```

```{r}
library(tidyverse)
library(spotifyr)
library(flexdashboard)
library(plotly)
library(compmus)

df <- get_playlist_audio_features("", "7KtmJXjsDmw8WEdtyL28L3?si=bc8137fcca564e11")

conditions <- c("QUARTIERS NORD", "Les dernières volontés de Mozart (Symphony)", "Cœur blanc", "Deux frères", 
                "L'EMPIRE DE MÉROÉ", "Dans la légende", "Capo Dei Capi Vol. II & III")
df$language <- ifelse(df$track.album.name %in% conditions, "French", "Dutch")

df_dutch <- df[df$language == "Dutch", ]
df_french <- df[df$language == "French", ]
 
cm23 <-
  bind_rows( 
    df_french |> mutate(category = "French"),
    df_dutch |> mutate(category = "Dutch")
  )
```

<!-- # Dutch Tracks -->
<!-- ```{r} -->
<!-- valueBox(nrow(df_dutch), icon="fa-music") -->
<!-- ``` -->

<!-- # French Tracks -->
<!-- ```{r} -->
<!-- valueBox(nrow(df_french), icon="fa-music") -->
<!-- ``` -->

Feature analysis {.storyboard} 
=========================================
### The Energetic Beat: Exploring the Relationship between Tempo and Danceability in Hip Hop Music 
```{r, fig.width=8, fig.height=7}
p1 <- cm23 %>%
  mutate(
    mode = ifelse(mode == 0, "Minor", "Major")
  ) %>%
  ggplot(
    aes(
      x = tempo, 
      y = danceability,
      size = energy,
      colour = mode
    )
  ) +
  geom_point() +
  geom_rug(linewidth = 0.2) +
  facet_wrap(~ category) +
  scale_x_continuous(
    limits = c(50, 200),
    breaks = c(50, 125, 200),
    minor_breaks = NULL
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = c(0, 0.50, 1),
    minor_breaks = NULL
  ) +
  scale_colour_brewer(
    type = "qual",
    palette = "Dark2"
  ) +
  scale_size_continuous(
    trans = "exp",
    #guide = "none",
    range = c(0, 2)
  ) + 
  theme_bw() +
  labs(
    x = "Tempo",
    y = "Danceability",
    colour = "Mode"
  ) + ggtitle("Relationship between tempo, danceability, energy in French & Dutch Hip hop")

ggplotly(p1)
```

***

In general, hip hop music is known for having a strong beat and a high level of energy. This is reflected in the tempo and danceability of many hip hop tracks, which are often fast and rhythmically complex.   
This plot really shows this relationship. A large proportion of tracks correspond to high energy and the "danceability" is well over 0.5 on average. Furthermore, we see interesting differences such as that the average tempo of Dutch raps is lower than French raps. The tempo distribution of French raps is greater than that of Dutch raps. This may be the reason why French raps more often give me more energy than Dutch raps.  

### Rhythmic Resonance bewteen Valence, Energy, and Loudness in French and Dutch Raps
```{r, fig.width=8, fig.height=7}
ggplot(cm23, mapping = aes(x = valence, y = energy, size=loudness)) +
  geom_boxplot(alpha = 0.5) + 
  geom_jitter(alpha = 0.5, color = "black") + 
  facet_wrap(~ category) +
  scale_size_continuous(
    trans = "exp",
    guide = "none",
    range = c(1, 5)) +
  ggtitle("Relationship between valence, energy and loudness in French & Dutch Hip hop") +
  theme_bw() 
```

***

This plot shows the relationship between the features "valence", "energy" and "Loudness" in French and Dutch raps. In the figure, the higher the valence and tempo be, the higher the loudness of the tracks be. Furthermore, the average energy of Dutch raps is lower, but whether this is significantly lower is questionable, as a discussion point is that the corpus contains slightly less Dutch rap than French rap. Lastly, the distribution of valence is almost the same in both cases


### Histograms of key modes for French and Dutch tracks in the corpus
```{r, fig.width=8, fig.height=7}
histcm23mode <- cm23 |>
     ggplot(aes(x = key_mode)) +
     geom_histogram(stat="count") +
     facet_wrap(~category) +
     ggtitle("The key mode the track is in for French & Dutch Hip hop") +
     theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplotly(histcm23mode)
```

***
This plot shows the key names and key modes the track is in for French and Dutch tracks.

### Use variation (standard deviation) in loudness, tempo, or durations as a variable in one of your visualisations.
```{r, fig.width=8, fig.height=7}
sliced_french <- df_french |> slice(1:30) |> add_audio_analysis()
sliced_dutch <- df_dutch |> slice(1:30) |> add_audio_analysis()

sliced_df <-
  sliced_french |>
  mutate(genre = "French") |>
  bind_rows(sliced_dutch |> mutate(genre = "Dutch"))

# Plot
sliced_df |>
  mutate(
    sections =
      map(
        sections,                                    # sections or segments
        summarise_at,
        vars(tempo, loudness, duration),             # features of interest
        list(section_mean = mean, section_sd = sd)   # aggregation functions
      )
  ) |>
  unnest(sections) |>
  ggplot(
    aes(
      x = tempo,
      y = tempo_section_sd,
      colour = genre,
      alpha = loudness
    )
  ) +
  geom_point(aes(size = duration / 60)) +
  geom_rug() +
  theme_minimal() +
  ylim(0, 5) +
  labs(
    x = "Mean Tempo (bpm)",
    y = "SD Tempo",
    colour = "Genre",
    size = "Duration (min)",
    alpha = "Volume (dBFS)"
  ) +
  scale_size_continuous(
    trans = "exp",
    range = c(1, 4))
```

***
...

Chromagram  
=========================================
Column 
-------------------------------------
***
In the case of French rap music in the corpus, analyzing the chromagram can provide insights into the tonal characteristics of the music and how they differ from other genres or regions. For example, it might reveal that French rappers tend to use certain chords or tonal centers more frequently than rappers from other countries, or that they use different melodic patterns or scales.

Analyzing the chromagram of a track of French rap such as this one could also be useful for identifying commonalities and differences between individual artists or sub-genres within French rap. For instance, it could reveal that one artist tends to use a particular tonal center or scale more frequently than others, or that certain sub-genres of French rap have distinct melodic or harmonic characteristics.

Comparing the chromagrams of two French tracks (outliers), "INTRO" and "MAINTENANT" from the playlist reveals significant differences in their pitch distribution. In "INTRO", over time, the focus of the pitches is on different Chords such as C, D and G, whereas in "MAINTENANT", there is a strong concentration of pitches around chord B and G. This means that "INTRO" is more complex and it has more different chord progressions. 

Column 
-------------------------------------
### "Chromagram of a recording of GIMS; INTRO." (Lowest danceability with one of the highest tempo)
```{r, fig.width=15, fig.height=7}
 wood <-
   get_tidy_audio_analysis("6H8njIysTyNsjD5CyPtqwN") |>
   select(segments) |>
   unnest(segments) |>
   select(start, duration, pitches)

 wood |>
   mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
   compmus_gather_chroma() |> 
   ggplot(
     aes(
       x = start + duration / 2,
       width = duration,
       y = pitch_class,
       fill = value
     )
   ) +
   geom_tile() +
   labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
   theme_minimal() +
   scale_fill_viridis_c()
```

### "Chromagram of a recording of GIMS; MAINTENANT" (Highest energy with averaged valence)
```{r, fig.width=15, fig.height=7}
 wood <-
   get_tidy_audio_analysis("5y2WMZ9UAYsA7qxRENID6l") |>
   select(segments) |>
   unnest(segments) |>
   select(start, duration, pitches)

 wood |>
   mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
   compmus_gather_chroma() |> 
   ggplot(
     aes(
       x = start + duration / 2,
       width = duration,
       y = pitch_class,
       fill = value
     )
   ) +
   geom_tile() +
   labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
   theme_minimal() +
   scale_fill_viridis_c()
```


Cepstrogram 
=========================================
Column
-------------------------------------
A cepstrogram serves as a graphical representation of a sound signal's spectral envelope. It displays time on the horizontal axis and frequency on the vertical axis, and is obtained by analyzing the cepstrum, which results from the Fourier transform of a sound's power spectrum logarithm. With its ability to analyze harmonics and other frequency components, the cepstrogram is a valuable tool for identifying the pitch and timbral distribution characteristics of individual notes or chords in musical recordings.

Comparing the tracks which I like the most reveals difference in between their timbral distributions. One of the most liked French and Dutch rap is "Au DD" of PNL and "Treinstation" of Boef respectively. 
On the one hand, in "Au DD", there isn't significantly a strong concentration (magnitude) of timbral characteristics around a specific spectral range. So, there is no frequency range where the sound is dominated by a particular quality or color that the song has a more balanced and consistent sound overall, without any one aspect of the sound dominating or standing out too much. 
On the other hand, "Treinstation" has more variation in different frequency ranges , with a stronger concentration of timbral characteristics around a certain frequency range. This could create a more unique and interesting sound for this piece. 

Column
-------------------------------------
### "Cepstrogram of a recording of PNL; Au DD" (French track also in "liked Songs" playlist)
```{r, fig.width=15, fig.height=7}
bztfrench <-
  get_tidy_audio_analysis("115Hll8WWkQLeiDyXpgr47?si=b0fe1f7f1b304fc7") |> # Change URI.
  compmus_align(bars, segments) |>                     # Change `bars`
  select(bars) |>                                      #   in all three
  unnest(bars) |>                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) |>
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  )

bztfrench |>
  compmus_gather_timbre() |>
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = basis,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
  scale_fill_viridis_c() +                              
  theme_classic()
```

### "Cepstrogram of a recording of Boef; Treinstation" (Dutch track also in "liked Songs" playlist)
```{r, fig.width=15, fig.height=7}
bztdutch <-
  get_tidy_audio_analysis("7yHf12Fxf8I1Q8Lhq6XqpZ?si=75225860823b4bfb") |> # Change URI.
  compmus_align(bars, segments) |>                     # Change `bars`
  select(bars) |>                                      #   in all three
  unnest(bars) |>                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) |>
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  )

bztdutch |>
  compmus_gather_timbre() |>
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = basis,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
  scale_fill_viridis_c() +                              
  theme_classic()
```



Self-similarity Matrices 
=========================================
Column
-------------------------------------
A self-similarity matrix helps to visualize and analyze the structure of a musical recording. It is created by comparing different sections of a piece of music and measuring the degree of similarity between them.

Here, both the similarity matrices based on chroma and timbre were evaluated for Au DD" of PNL and "Treinstation" of Boef.
The timbre-based self-similarity matrix (which refers to the tone color or quality of the sound) for "Au DD" doesn't clearly show block which indicate repeated patterns in the sound, whereas the self-similarity matrix based on timbre for "Treinstation" does. However, if one look carefully at the timbre-based self-similarity matrix for "Au DD", one can see two small lines parallel to the main diagonal line in the figure which means that there are repetitions in the track. 

The chroma-based self-similarity matrix (which refers to the pitch content of the music) of a recording of Boef; Treinstation” shows clearly small parallel lines indicating repeated patterns in the sound, whereas The chroma-based self-similarity matrix for "Au DD" does not.  

Column
-------------------------------------
### "Self-similarity Matrix (Timbre) of a recording of PNL; Au DD" (French)
```{r, fig.width=7, fig.height=7}
bztfrench |>
  compmus_self_similarity(timbre, "cosine") |> 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")

```

### "Self-similarity Matrix (Timbre) of a recording of Boef; Treinstation" (Dutch)
```{r, fig.width=7, fig.height=7}
bztdutch |>
  compmus_self_similarity(timbre, "cosine") |> 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")

```

Column
-------------------------------------
### "Self-similarity Matrix (Chroma) of a recording of PNL; Au DD"
```{r, fig.width=7, fig.height=7}
bztfrench |>
  compmus_self_similarity(pitches, "cosine") |> 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")

```

### "Self-similarity Matrix (Chroma) of a recording of Boef; Treinstation" (Dutch)
```{r, fig.width=7, fig.height=7}
bztdutch |>
  compmus_self_similarity(pitches, "cosine") |> 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")

```

Keygram 
=========================================
Column
-------------------------------------
the differences and similarities between French and Dutch rap music were explored through a collection of 300 tracks by French and Dutch rap artists. The portfolio found that French rap is characterized by its strong tradition of chanson and the incorporation of African and Caribbean music, while Dutch rap incorporates elements of gabber and house. Both groups use a fast and energetic flow, but there are differences in tempo and energy levels. 

Column
-------------------------------------
### "Keygram of a recording of GIMS; Apres Vous Madame" (French track also in "liked Songs" playlist)
```{r, fig.width=15, fig.height=7}
circshift <- function(v, n) {
  if (n == 0) v else c(tail(v, n), head(v, -n))
}

#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )


Keygram <-
  get_tidy_audio_analysis("1kVQzCxcSVhqf4rLJyGk50?si=d11761254d384f40") |>
  compmus_align(sections, segments) |>
  select(sections) |>
  unnest(sections) |>
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "mean", norm = "manhattan"
      )
  )

Keygram |> 
  compmus_match_pitch_template(
    key_templates,         # Change to chord_templates if descired
    method = "euclidean",  # Try different distance metrics
    norm = "manhattan"     # Try different norms
  ) |>
  ggplot(
    aes(x = start + duration / 2, width = duration, y = name, fill = d)
  ) +
  geom_tile() +
  scale_fill_viridis_c(guide = "none") +
  theme_minimal() +
  labs(x = "Time (s)", y = "")

```

Conclusion 
=========================================
In this portfolio, the differences and similarities between French and Dutch rap music were explored through a collection of 300 tracks by French and Dutch rap artists. The portfolio found that French rap is characterized by its strong tradition of chanson and the incorporation of African and Caribbean music, while Dutch rap incorporates elements of gabber and house. Both groups use a fast and energetic flow, but there are differences in tempo and energy levels. The portfolio also analyzed the chromagrams of individual tracks, revealing differences in pitch distribution and chord progressions. Overall, the portfolio provides insight into the unique characteristics of French and Dutch rap and how they contribute to the development of hip-hop as a global cultural form.


